### 2022-02-22 paper/output/document/thesis_5_discussion.tex
cleaning chapter 5

The implementation of this paper was inspired by and builds on Benzmüller, Parent, and Farjami's 
foundational work with the LogiKEy framework for machine ethics, which includes their implementation 
of DDL in Isabelle \citep{BFP, logikey}. The LogiKEy project has been used to study metaphysics 
\citep{godel, metaphysics1}, law \citep{constitution}, and ethics \citep{gewirth}, but not 
Kant's categorical imperative.%

@@

\isadelimdocument
\endisadelimdocument
\isatagdocument
\isamarkupsubsection{Conclusion%
}
\isamarkuptrue%
\endisatagdocument
{\isafolddocument}%
\isadelimdocument
\endisadelimdocument

@@

>>>>

The implementation of this paper was inspired by and builds on Benzmüller, Parent, and Farjami's 
foundational work with the LogiKEy framework for machine ethics, which includes their implementation 
of DDL in Isabelle \citep{BFP, logikey}. The LogiKEy project has been used to study metaphysics 
\citep{godel, metaphysics1}, law \citep{constitution}, and ethics \citep{gewirth}, but not 
Kant's categorical imperative.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Conclusion%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this thesis, I present a proof-of-concept implementation of automated Kantian ethics. My system
takes as input a potential action, appropriately represented, and can prove that it is obligated, 
prohibited or permissible. I represent Kant's Formula of Universal Law in a deontic logic and 
implement this logic in the Isabelle/HOL interactive theorem prover, which can automatically prove or 
refute theorems in my custom logic. I also contribute a testing framework that I use to demonstrate that
my implementation of Kantian ethics is more faithful to philosophical literature than two other 
potential implementations. My completed system can, when given appropriate factual background, make
philosophically mature judgements about complex moral dilemmas. This work is one step towards building 
morally sophisticated artifical agents.

The idea of fully automated artificial agents navigating the world without human supervision
may be terrifying, but progress in AI indicates that such a future is likely closer than we think. 
Philosophers, regulators, and computer scientists are sounding the alarm about the dangers of developing
this kind of AI. Insofar as developers will continue to ignore these warnings and develop increasingly 
independent AI, there is a dire need to program such AI with some notion of ethics. If AI is navigating
human society, then it is making decisions with an ethical component at all times. Ethics is inescapable; 
if AI developers and computer scientists ignore it, then they will be building machines that make decisions
based on some set of unknown, implicit ethical values. Countless examples, from the Alleghany family services screening
algorithm that is biased against poor families to search algorithms that associate black-sounding names
with crime, demonstrate that such implicit ethics simply codifies the biases, prejudices, and moral
failings of the society in which it is developed. AI will inevitably make judgements on moral dilemmas, 
and automated ethics is necessary to make these judgements morally correct.

Given that the discpline of philosophy has spent centuries debating such judgements and their theoretical 
underpinnings, such AI will be most trustworthy, nuanced, consistent, and mature when it is faithful to 
philosophical literature. In order to develop high-quality automated ethics, computer scientists and 
philosophers must work together. This thesis is an experiment in marrying philosophy and computer science 
to create automated ethics that is both technically and philosophically advanced. Neither discipline alone
can address the pressing need for ethical AI.

This work is an early proof-of-concept. It demonstrates the potential of top-down, logic programming 
approaches to automated ethics and shows that it is possible to faithfully automate an ethical theory as
complex as Kantian ethics. There are open questions that must be resolved before a system like this 
could be used in practice, but this project demonstrates that these questions are within closer reach
than they may seem. Automated ethics does not need to limit itself to simple, flattened versions of 
ethical theories. With technical and philosophical progress, faithful automated ethics is possible.
Growing public consciousness about the dangers of unregulated AI is creating momentum in machine ethics; 
work like Delphi demonstrates that the time is ripe to create usable, reliable automated ethics. This 
thesis is one step towards building computers that can think ethically in the richest sense of the word.%

@@

@@

%:%719=590%:%
%:%720=591%:%
%:%721=592%:%
%:%722=593%:%
%:%723=594%:%
%:%724=595%:%
%:%725=596%:%
%:%726=597%:%
%:%727=598%:%
%:%728=599%:%
%:%729=600%:%
%:%730=601%:%
%:%731=602%:%
%:%732=603%:%
%:%733=604%:%
%:%734=605%:%
%:%735=606%:%
%:%736=607%:%
%:%737=608%:%
%:%738=609%:%
%:%739=610%:%
%:%740=611%:%
%:%741=612%:%
%:%742=613%:%
%:%743=614%:%
%:%744=615%:%
%:%745=616%:%
%:%746=617%:%
%:%747=618%:%
%:%748=619%:%
%:%749=620%:%
%:%750=621%:%
%:%751=622%:%
%:%752=623%:%
%:%753=624%:%
%:%754=625%:%
%:%755=626%:%
%:%756=627%:%
%:%757=628%:%
